Finite Time Analysis of Linear Stochastic Approximation

### 2017-05-05 intro.tex
updated intro

Stochastic Approximation (SA) algorithms can be expressed in the general form
\begin{align}\label{eq:sa}
\theta_t=\theta_{t-1}+\alpha_t f(\theta_{t-1},\zeta_t),
where $\theta_t\in \R^d$ are the iterates, and $f(\theta_{t-1},\zeta_t)$ of the function $g(\theta_{t-1})=\EE{f(\theta_{t-1})}_{\zeta_t}$, where the expectation is with respect to some noise process $\zeta_t$. At each time step, the algorithm updates the iterates in increments proportional to a step-size $\alpha_t>0$ and the idea is to eventually converge to $\ts\in \R^d$ which is a desired optimum or a fixed point. Due to their inherent online nature SA algorithms appear widely in machine learning applications \cite{konda,bhatnagar,borkar,gtd,td,bach}, where we have to learn from a noisy data stream.\par
The choice of the step-size $\alpha_t$ needs to trade-off learning rate versus filtering noise, so to speak. In general, one can guess that larger step-size lead to faster learning and at the same time smaller step-sizes are better to handle noise. A popular choice of step-size schedule that is known handle the noise $\alpha_t=O(\frac{1}{t})$. However, since the step size are diminishing, it can hurt the learning rate. An alternative suggestion is to use more aggressive step sizes without minding the noise in the iterates $\theta_t$ in \eqref{eq:sa}, however, introduce additional averaging (known as Polyak-Rupert averaging) to filter the noise. The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$, and an analysis of the PR-averaging technique is known from \cite{}.\par

@@

\item We also show (via counter example) that not all classes of general LSAs `admit' a constant uniform step size. Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.

@@

>>>>

Stochastic Approximation (SA) algorithms can be expressed in the general form
\begin{align}\label{eq:sa}
\theta_t=\theta_{t-1}+\alpha_t f(\theta_{t-1},\zeta_t),
\end{align}
where $\theta_t\in \R^d$ are the iterates, and $f(\theta_{t-1},\zeta_t)$ of the function $g(\theta_{t-1})=\EE{f(\theta_{t-1})}_{\zeta_t}$, where the expectation is with respect to some noise process $\zeta_t$. At each time step, the algorithm updates the iterates in increments proportional to a step-size $\alpha_t>0$ and the idea is to eventually converge to $\ts\in \R^d$ which is a desired optimum or a fixed point. Due to their inherent online nature SA algorithms appear widely in machine learning applications \cite{konda,bhatnagar,borkar,gtd,td,bach}, where we have to learn from a noisy data stream.\par
The choice of the step-size $\alpha_t$ needs to trade-off learning rate versus filtering noise, so to speak. In general, one can guess that larger step-size lead to faster learning and at the same time smaller step-sizes are better to handle noise. A popular choice of step-size schedule that is known handle the noise $\alpha_t=O(\frac{1}{t})$. However, since the step size are diminishing, it can hurt the learning rate. An alternative suggestion is to use more aggressive step sizes without minding the noise in the iterates $\theta_t$ in \eqref{eq:sa}, however, introduce additional averaging (known as Polyak-Rupert averaging) to filter the noise. The averaged iterates can be given by $\th_t\eqdef\frac{1}{t}\sum_{s=0}^t \theta_t$, and an analysis of the PR-averaging technique is known from \cite{}.\par

@@

\item We also show (via counter example) that not all classes of general LSAs `admit' a constant step-size that is uniform across all the problems in the class. Further, we also argue that the subclass of LSAs related to linear prediction problem does `admit' a problem independent constant step size.
\item The assumptions in our setting is comparable to those in the asymptotic analysis based on ordinary differential equation (ODE) method by \cite{borkarmeyn,borkarbook}.
\item Two timescale LSA algorithms are those that use two set of iterates (namely slower and faster) and two separate step-size schedules. Our results also hold for two timescale LSAs as long as the two different step-sizes are held constant with respect to time. Thus our setting not only holds for TD(0) but also extends to GTD/GTD2 and TDC.
\item On an intuitive level, the error dynamics of the constant step-size plus PR-averaging procedure is similar to the behavior of partial sums of a geometric series. This coupled with the fact that the result also encompasses two timescale LSAs presents a simple and unified view of the various TD algorithms.
\item The break up of error into bias and variance terms and their respective rates were previously worked out by \cite{bach} in the linear prediction setting, and here we extend the same to the general class of LSAs. Nevertheless, there are certain key differences between the general LSA setting that we consider to the setting considered by \cite{bach}, which are listed as under.

@@

\begin{comment}
